## Final list of items

```{r, include=FALSE}
set.seed(2137) # set seed for replicability
```

Lastly, we need to generate the final list of items for further empirical testing:

* items included based on the identified factor structure
* `SOR` items included based on theoretical / conceptual grounds.

### Load inputs

We load data required for the analysis, together with some helper variables (`code_to_pl`, `code_to_en`).

``` {r}
load("./03/input/dataset.RData")
```

Furthermore, we load information about which items we want to include to the final list of items. This information is a result of qualitative assessment, that was performed by the team members.

``` {r}
items7 = read.table("./04/input/items-per-factor-dropouts-7.tsv", header = T, sep = "\t",  quote = "", encoding = "UTF-8")
rownames(items7) = items7$CODE

items11 = read.table("./04/input/items-per-factor-dropouts-11.tsv", header = T, sep = "\t",  quote = "", encoding = "UTF-8")
rownames(items11) = items11$CODE
```

``` {r, include=FALSE}
selected7 = is.na(items7$LOW) & is.na(items7$SIM) & is.na(items7$UNC) & is.na(items7$EMO)
selected11 = is.na(items11$LOW) & is.na(items11$SIM) & is.na(items11$UNC) & is.na(items11$EMO)
```

Overall we decided to keep:

* `r sum(selected7)` items, identified based on the 7 factor model
* `r sum(selected11)` `SOR` items, identified based on the 11 factor model.

### Prepare outputs

``` {r}
ITEMS =  list(items7, items11)
output = data.frame(matrix(ncol = 3, nrow = 0))

for (i in 1:2) {
  
  items = ITEMS[[i]]
  rownames(items) = items$CODE
  
  selected = is.na(items$LOW) & is.na(items$SIM) & is.na(items$UNC) & is.na(items$EMO)

  codes = items$CODE[selected]
  col1 = as.character(code_to_pl[codes])
  col2 = as.character(code_to_en[codes])
  col3 = as.character(codes)
  
  out = as.data.frame(cbind(col1, col2, col3))
  output = rbind(output,out)
}
```

### Save outputs

``` {r}
write.table(output, file="./05/output/questionnaire.tsv", row.names = F, col.names = F, sep='\t', fileEncoding = "UTF-8")
```